Draft Article

### 2019-11-08 article.tex
starting glazer theory

Under traditional methods of analysis, espionage certainly seems like something that states should be doing more to discourage. Then results of intelligence operations can have devastating, even existential effects on the targeted nation's security. One of the most traditional form of espionage is stealing military secrets from a foreign power, and in that arena China has been relentless. A US Navy report from March 2019 declared Chinese intelligence operations so extensive that they had substantively altered the balance of power between the two states.\footcite{lubold_navy_2019} ``Long-term, US future military advantage is being diminished by years of IP exfiltration from the DoD, DoN, and DIB,'' the report read, ``all with little to no adverse consequences to the thieves.''\footcite[p.~6]{bayer_cybersecurity_2019} The Navy's forecasts were dire: ``If the current trend continues unimpeded, the US will soon lose its status as the dominant global economic power.''\footcite[p.~5]{bayer_cybersecurity_2019} The Navy is not a disinterested party, of course, and there are reasons to take its conclusions with a grain of salt; modern military technology is remarkably complex and China cannot simply steal its way to military dominance.\footnote{For a good discussion of these limitations, see \cite{gilli_why_2019}} Nonetheless, the possibilities are alarming and the consequences for remain minimal.

Not only does espionage have potentially disastrous consequences but the system is set up so that the structure of civilian intelligence services is never fundamentally threatened. The low percentage of responses for cyber-espionage have their basis in a similarly low response rate for traditional espionage. The traditional punishment for espionage is to declare certain embassy officials \emph{persona non grata}, which often results in ``tit-for-tat'' diplomatic expulsions. The name is taken from a highly-successful strategy for the prisoner's dilemma wherein the player simply mimics the choice that that their opponent made in the last round. By employing the tit-for-tat strategy, the player never gets taken advantage of for more than one turn and can always return to cooperation if their opponent unilaterally signals a willingness to do the same.

@@

Therefore, I propose that an iterated model can describe how states cooperate to preserve the liminal legal status of espionage. My model changes the parameters from above so that the decision to cooperate or defect is based not in the amount of punishment received (essentially measured by the number of diplomats expelled), but rather the type. Each time a state is forced to respond to a successful intelligence operation by a rival---such as the reveal of a high-level spy or a major data breach---it can either limit its response to that which is traditionally associated with espionage, or it can impose unprecedented costs that might discourage such an operation in the future. Table \ref{espionage-matrix} describes a single instance of this choice. States signal their interest in cooperating by limiting their retaliation in response to espionage, no matter how damaging the operation is to their national security. Each instance is a chance to play the game with fresh information; each espionage case is an opportunity to cooperate, reaffirming the norm that states always respond to espionage in a specific way, or defect, adding an additional cost to the attempt.

% https://tex.stackexchange.com/questions/249480/creating-a-payoff-matrix-using-latex-tabular-environment
\begin{table}[ht]
\centering
\setlength{\extrarowheight}{2pt}
\small
\begin{tabular}{cc|c|c|}
  & \multicolumn{1}{c}{} & \multicolumn{2}{c}{USSR}\\
  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{Tolerate Espionage}  & \multicolumn{1}{c}{Punish Espionage} \\\cline{3-4}
  \multirow{3}*{USA}  & Tolerate Espionage & \makecell{~\\Both sides gain intelligence \\~} & \makecell{~\\ Only the USSR continues spying \\ ~} \\\cline{3-4}
  & Punish Espionage & \makecell{~\\ Only the US continues spying \\~} & \makecell{~\\ No intelligence either direction \\~} \\\cline{3-4}
\end{tabular}
\caption{Espionage consequences decision matrix}
\label{espionage-matrix}
\end{table}

The purpose of analyzing espionage norms this way is to remove the mystique.

Even though the US and the USSR have the opportunity to defect by imposing harsher punishments in response to espionage, the gains would be short-lived, because the other state is guaranteed to immediately do the same---tit-for-tat. Espionage consequences are an iterated game that both parties are playing the exactly the same way, so the only two possible outcomes long-term are mutual defection or mutual cooperation.

Given those two options, it makes intuitive sense that in the case of Cold War espionage both the US and the USSR chose to cooperate. Both sides gained a great deal from their intelligence operations; Soviet intelligence had an easier time operating in Washington than the US did in the USSR, but American intelligence was still formidable and Soviet defections were frequent.\footcite[p.~380]{macrakis_technophilic_2010} When the relative capabilities are not too lopsided, there is a pleasing reciprocity to espionage that makes cooperation easier.

>>>>

Under traditional methods of analysis, espionage certainly seems like something that states should be doing more to discourage. Then results of intelligence operations can have devastating, even existential effects on the targeted nation's security. Arguably the most traditional form of espionage is stealing military secrets from a foreign power, and in that arena China has been relentless. A US Navy report from March 2019 declared Chinese intelligence operations so extensive that they had substantively altered the balance of power between the two states.\footcite{lubold_navy_2019} ``Long-term, US future military advantage is being diminished by years of IP exfiltration from the DoD, DoN, and DIB,'' the report read, ``all with little to no adverse consequences to the thieves.''\footcite[p.~6]{bayer_cybersecurity_2019} The Navy's forecasts were dire: ``If the current trend continues unimpeded, the US will soon lose its status as the dominant global economic power.''\footcite[p.~5]{bayer_cybersecurity_2019} The Navy is not a disinterested party, of course, and there are reasons to take its conclusions with a grain of salt; modern military technology is remarkably complex and China cannot simply steal its way to military dominance.\footnote{For a good discussion of these limitations, see \cite{gilli_why_2019}} Nonetheless, the opportunities for harming national security are alarming and the consequences for doing so remain minimal.

Not only does espionage have potentially disastrous consequences but steps that might actually curb it are out of the question. The system is set up so that the structure of civilian intelligence services is never fundamentally threatened; the minimal responses for cyber-espionage have their roots in the similarly minimal responses for classical espionage. The traditional punishment for espionage is to declare certain embassy officials \emph{persona non grata}, which often results in ``tit-for-tat'' diplomatic expulsions. The name is taken from a highly-successful strategy for the prisoner's dilemma wherein the player simply mimics the choice that that their opponent made in the last round. By employing the tit-for-tat strategy, the player never gets taken advantage of for more than one turn and can always return to cooperation if their opponent unilaterally signals a willingness to do the same.

@@

% https://tex.stackexchange.com/questions/249480/creating-a-payoff-matrix-using-latex-tabular-environment
\begin{table}[ht]
\centering
\setlength{\extrarowheight}{2pt}
\small
\begin{tabular}{cc|c|c|}
  & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Country 2}\\
  & \multicolumn{1}{c}{} & \multicolumn{1}{c}{Tolerate Espionage}  & \multicolumn{1}{c}{Punish Espionage} \\\cline{3-4}
  \multirow{3}*{Country 1}  & Tolerate Espionage & \makecell{~\\Both sides gain intelligence \\~} & \makecell{~\\ Only the USSR continues spying \\ ~} \\\cline{3-4}
  & Punish Espionage & \makecell{~\\ Only the US continues spying \\~} & \makecell{~\\ No intelligence either direction \\~} \\\cline{3-4}
\end{tabular}
\caption{Espionage consequences decision matrix}
\label{espionage-matrix}
\end{table}

Therefore, I propose an iterated model to describe how states cooperate to preserve the liminal legal status of espionage. My model changes the parameters so that the decision to cooperate or defect is based not in the amount of punishment received, but rather the type. Each time a state is forced to respond to a successful intelligence operation by a rival---such as the reveal of a high-level spy or a major data breach---it can either limit its response to one which is traditionally associated with espionage, or it can impose unprecedented costs that might discourage such an operation in the future. Table \ref{espionage-matrix} describes a single instance of this choice. States signal their interest in cooperating by limiting their retaliation in response to espionage, no matter how damaging the operation is to their national security. Each instance is a chance to play the game with fresh information; each espionage case is an opportunity to cooperate, reaffirming the norm that states always respond to espionage in a specific way, or defect, adding an additional cost to the attempt.

Through this model two important things become clear. First, even though either country has the opportunity to defect by imposing harsher punishments in response to espionage, its gains would be short-lived, because the other country is guaranteed to immediately do the same---tit-for-tat. Espionage consequences are an iterated game that both parties are playing the exact same way, so, in the long run, the only two possible outcomes are mutual defection or mutual cooperation. In the realm of traditional espionage, we've only ever seen mutual cooperation, but the key observation is that any country has the power to change that by defecting. Even if the defector's rival would have preferred to maintain the status quo, the rival has no option other than to defect as well, or else be left at a senseless disadvantage. The relatively open flow of information is only possible with a universal buy-in.

Second, while there is unquestionably a cultural norm informing the response to intelligence operations, that norm is doing a lot of work to maintain the status quo. So much work, in fact, that tradition could not possibly explain the entire phenomenon. Stripped of its mystique, the benefits of allowing intelligence operations to continue receiving light treatment are, like any other state decision, a simple cost-benefit analysis. And therefore, regardless of its correctness or actual explanatory power, there must be \emph{some} rational explanation that political leaders employ when they decide that preserving their ability to spy is worth suffering through the inevitable leaks that result from espionage conducted against them.

\subsection{Explaining norms cooperation}

% Given those two options, it makes intuitive sense that in the case of Cold War espionage both the US and the USSR chose to cooperate. Both sides gained a great deal from their intelligence operations; Soviet intelligence had an easier time operating in Washington than the US did in the USSR, but American intelligence was still formidable and Soviet defections were frequent.\footcite[p.~380]{macrakis_technophilic_2010} When the relative capabilities are not too lopsided, there is a pleasing reciprocity to espionage that makes cooperation easier.
